Put Title here

### 2018-12-19 General WLD combinatorics.tex
Made changes to section 2.3, stopping after 2.14

[outline of section, postponed until section structure is finalised]

The main results of this section are the following: we show that two admissible Wilson loop diagrams define the same matroid if and only if they are equivalent (Theorem \ref{same matroid iff equiv}), and we obtain a formula for the number of admissible Wilson loop diagrams in each equivalence class (Corollary \ref{number of equiv diagrams}).  [sentence about why this matters]

\subsection{Wilson loop diagrams as matroids}\label{sec matroid background}

@@

\begin{lem}\label{lem facts about WLD matroids}
Let $W = (\cP,n)$ be an admissible Wilson loop diagram. Then:
\begin{enumerate}
\item The rank of a set $V \subseteq [n]$ is bounded above by $\min\{|V|,|\Prop(V)|\}$, with $\rk(V) = |V|$ if and only if $V$ is an independent set.
%\item Let $v \in [n]$ and $q \in \Prop(v)$. Then for any $V \subset [n]$ such that $q \not \in \Prop(V)$, we have $\rk(V \cup \{v\}) = \rk(V) +1$, i.e. adding a vertex that supports a new propagator to a set increases the rank of the set.
\item If $C \subseteq [n]$ is a cycle, then $\rk(C) = |Prop(C)|$.
\item If $[n]$ can be partitioned into at least two non-empty sets, each of which support different sets of propagators that form a partition of the propagotor set, \bas [n] = \sqcup_i V(P_i) \quad \textrm{s.t.} \quad \sqcup P_i = \cP; \quad V(P_i) \cap V(P_j) = \emptyset; \quad  P_i \cap P_j = \emptyset \;, \eas then the matroid $M(W)$ is seperable, \bas M(W) = \bigoplus_i M(P_i, V(P_i)) \;.\eas 
\end{enumerate}
\end{lem}
\begin{proof}
The first part of (1) is \cite[Equation (9)]{wilsonloop} and surrounding discussion, and the second part is standard matroid theory. (2) is \cite[Lemma 3.27]{wilsonloop}. (3) is a direct consequence of \cite[Lemma 3.20]{wilsonloop} and the fact that $F(P_1)^c = V(P_1^c)$
\end{proof}

@@

\subsection{Polygon partitions of Wilson loop diagrams}

@@

\subsection{Matroid properties of exact subdiagrams}

Since Corollary \ref{uniqueproppartitioncor} allows us to decompose any admissible Wilson loop diagram into a collection of maximal exact subdiagrams, in this section we examine the matroid properties of exact subdiagrams more closely. In particular, we show that the matroids associated to exact subdiagrams are uniform (Theorem \ref{??}) and that

@@

\begin{lem}\label{lem properties of prop flats} \sanote{do we need the first 2? They follow directly from the definition 1.2. Do we use them later?}
Let $F(P)$ be a propagator flat as defined above. Then
\begin{enumerate}
\item $F(\emptyset)$ is exactly the set of vertices which support no propagators, while for $P \neq \emptyset$ we have \[F(P) = \big(V(P) \setminus V(P^c)\big) \cup F(\emptyset).\] \note{updated (1) to match new definition of $F(P)$; }
\item If $Q\subseteq P$ then $F(Q) \subseteq F(P)$.
\item $F(P)$ is a flat of $M(W)$, thus justifying the name ``propagator flat''.
\end{enumerate}
\end{lem}
\begin{proof}
The proof of (1) and (2) are routine applications of the definition and are omitted.

@@

Let $v \in C$. Clearly $\Prop(v) \subseteq \Prop(C)$, and so by Lemma \ref{lem properties of prop flats} we have $F(\Prop(v)) \subseteq F(\Prop(C))$. Since $v \in F(\Prop(v))$ by the definition of propagator flat, we have $v \in F(\Prop(C))$ as required.

Now suppose there exists some $w \in  F(\Prop(C)) \setminus C$. Let $B$ be an independent subset of $C$ of maximal rank; we first show that $B \cup \{w\}$ is a dependent set in $M(W)$. Indeed, by \cite[Proposition 3.10]{wilsonloop}, there exists some $Q \subseteq \Prop(B)$ with $|Q| = |B|$ and such that $B$ is a basis in $M(Q,V(Q))$. We therefore have
\[Q \subseteq \Prop(B) \subseteq \Prop(C),\]
with $|\Prop(C)| = \rk(C)$ by Lemma \ref{lem facts about WLD matroids} and $|Q| = |B| = \rk(C)$ by the choice of $B$. Thus $Q = \Prop(C)$, and so $B$ is a basis in the matroid defined by $W' := (\Prop(C),V(\Prop(C)))$. All that remains is to note that $w \in F(\Prop(C)) \subseteq V(\Prop(C))$, i.e. $w$ is a vertex in $W'$; thus $B \cup\{w\}$ is dependent in $W'$, and hence also in $W$.  This also implies that $\rk(F\cup w) = \rk(F)$ since $B\subseteq F$ and so since $F$ is a flat we must have $w\in F$.

Returning our attention to $M(W)$, the dependent set $B \cup \{w\}$ must contain a circuit $C'$, with $w \in C'$ since $B$ was an independent set. Now $C \cup C' = C \cup \{w\} \supsetneq C$ is a cycle which, since $w\in F$, contradicts the maximality of $C$.

@@

For any $e \not\in F$, we certainly have $\rk((F\setminus C)\cup\{e\}) = \rk(F\setminus C) +1$ since $F$ is a flat. Now let $e \in C$, and suppose that $\rk((F\setminus C)\cup\{e\}) = \rk(F\setminus C)$. This implies that $(F\setminus C)\cup \{e\}$ is dependent, and hence contains a circuit. This circuit must contain at least two elements \note{What if $e \in C$ supported no propagators? $\{e\}$ would then be a circuit on its own}(since all vertices support at least one propagator), but this contradicts the fact that $C$ was the union of all circuits in $F$.

@@

\begin{thm} \label{exactcircuitlem}
Let $W':= (P, V(P))$ be a subdiagram of an admissible Wilson loop diagram $W= (\cP, [n])$. Then $W'$ is an exact subdiagram if and only if $M(W')$ is a uniform matroid of rank $|P|$.
\end{thm}

@@

    Taking all the possibilities for the left two cases of Figure~\ref{fig 3 cases}, the intervals $J^{(W)}_r(*)$ are still cyclic for all $r \neq p$, and we can assemble the intervals for the $J_p^{(W)}(*)$ as follows.
    \begin{itemize}
      \item 
    If $m=l$ then either $a=2$ (so $l+1=m+1$, $j=m+2$, and $j+1=m+3$) or $a=3$ (so $l+1=m+1$, $l+2=m+2$, $j=m+3$, and $j+1$ is not in the support of $p$), and in both cases
    \[
    J^{(W)}_p(m) = [m+4, m], \  J^{(W)}_p(m+1) = \{m+1\}, \  J^{(W)}_p(m+2) = \{m+2\}, \  J^{(W)}_p(m+3) = \{m+3\},
    \]
    which are nonempty and otherwise as required.
  \item
        If $m=i+1$ then checking each of the three different possibilities for $a$ we likewise get
    \[
    J^{(W)}_p(m) = [m+4, d-1], \  J^{(W)}_p(m+1) = [d, m+1], \  J^{(W)}_p(m+2) = \{m+2\}, \  J^{(W)}_p(m+3) = \{m+3\},
    \]
    which are nonempty and otherwise as required.
  \item If $m=i$ then $a=1$ or $a=2$, in the former case $l=m+2$, $j=m+3$ and $j+1$ is not in the support of $p$ so
    \[
    J^{(W)}_p(m) = \{m+3\}, \  J^{(W)}_p(m+1) = [m+4, d-1], \  J^{(W)}_p(m+2) = [d, m+1], \  J^{(W)}_p(m+3) = \{m+2\},
    \]
    while in the latter $l=m+2$, $l+1=m+3$, and $j$ and $j+1$ are not in the support of $p$ so
    \[
    J^{(W)}_p(m) = [m+4, j+1], \  J^{(W)}_p(m+1) = [j+2, d-1], \  J^{(W)}_p(m+2) = [d, m+1], \  J^{(W)}_p(m+3) = \{m+3\},
    \]
    which are again as required.
    \end{itemize}

>>>>

In subsection 2.1, we discuss a few matroidal facts about Wilson loop diagrams. In section 2.2, we show that there is a one to one correspondence between exact subdiagrams of Wilson loop diagrams and triangulated pieces of the corresponding polygon partion. In section 2.3, we return to some matroidal facts about Wilson loop diagrams, and exact subdiagrams in partuclar. We show that a subdiagram of $W$ defines a uniform matroid if and only if the subdiagram is exact. This is then used to prove the main result of section 2.3, which is that that two admissible Wilson loop diagrams define the same matroid if and only if they are equivalent (Theorem \ref{same matroid iff equiv}), and we obtain a formula for the number of admissible Wilson loop diagrams in each equivalence class (Corollary \ref{number of equiv diagrams}).  [sentence about why this matters] \todo{much of the above is just text. PUT REFERENCES IN}

\subsection{Wilson loop diagrams as matroids\label{sec matroid background}}

@@

\begin{lem}\label{lem facts about WLD matroids}
Let $W = (\cP,n)$ be an admissible Wilson loop diagram. Then:
\begin{enumerate}
\item The rank of a set $V \subseteq [n]$ is bounded above by $\min\{|V|,|\Prop(V)|\}$, with $\rk(V) = |V|$ if and only if $V$ is an independent set.
%\item Let $v \in [n]$ and $q \in \Prop(v)$. Then for any $V \subset [n]$ such that $q \not \in \Prop(V)$, we have $\rk(V \cup \{v\}) = \rk(V) +1$, i.e. adding a vertex that supports a new propagator to a set increases the rank of the set.
\item If $C \subseteq [n]$ is a cycle, then $\rk(C) = |Prop(C)|$.
\item If $[n]$ can be partitioned into at least two non-empty sets, each of which support different sets of propagators that form a partition of the propagotor set, \bas [n] = \sqcup_i V(P_i) \quad \textrm{s.t.} \quad \sqcup P_i = \cP; \quad V(P_i) \cap V(P_j) = \emptyset; \quad  P_i \cap P_j = \emptyset \;, \eas then the matroid $M(W)$ is seperable, \bas M(W) = \bigoplus_i M(P_i, V(P_i)) \;.\eas 
\item $F(P)$ is a flat of $M(W)$, thus justifying the name ``propagator flat''.
\end{enumerate}
\end{lem}
\begin{proof}
The first part of (1) is \cite[Equation (9)]{wilsonloop} and surrounding discussion, and the second part is standard matroid theory. (2) is \cite[Lemma 3.27]{wilsonloop}. (3) is a direct consequence of \cite[Lemma 3.20]{wilsonloop} and the fact that $F(P_1)^c = V(P_1^c)$

To prove (4), we need to show that $F(P)$ is maximally dependent. If $F(P) = [n]$ then this is automatic, so suppose not and let $v \in [n] \setminus F(P)$. In other words, $v \in V(P^c)$ and so $v$ supports some propagator $q \not\in P$.  Let $S \subseteq F(P)$ be an independent set of maximal size. Then $\Prop(S) \subseteq P$ by (1), and no subset of $S$ supports fewer propagators than the number of vertices it contains (this is the definition of an independent set in $M(W)$). Since $v$ supports a new propagator $q \not\in P$, the set $S \cup \{v\} \subseteq F(P) \cup\{v\}$ also satisfies this independence condition. Thus $\rk(F(P) \cup\{v\}) = \rk(F(P)) + 1$, as required.
\end{proof}

@@

\subsection{Polygon partitions of Wilson loop diagrams\label{sec: polygon partitions}}

@@

\subsection{Matroidal properties of exact subdiagrams \label{sec: exact diagram matroidal props}}

Since Corollary \ref{uniqueproppartitioncor} allows us to decompose any admissible Wilson loop diagram into a collection of maximal exact subdiagrams, in this section we examine the matroid properties of exact subdiagrams more closely. In particular, we show that given an exact subdiagram $(P, V(P))$, it can be written as the contraction of the matroid $M(W)$ by the complementary propagator flat, $F(P^c)$. Furthermore, if $(P, V(P))$ is a maximal exact subdiagrams, its complementary propagator flat is a cyclic flat. Finally, we show that matroids associated to exact subdiagrams are uniform (Theorem \ref{exactuniformthm}).

We begin by proving some useful facts about flats of matroids associated to admissible Wilson loop diagrams.

\begin{comment}

@@

\begin{lem}\label{lem properties of prop flats} \sanote{do we need the first 2? They follow directly from the definition 1.2. Do we use them later?}
Let $F(P)$ be a propagator flat as defined above. Then
\begin{enumerate}
\item $F(\emptyset)$ is exactly the set of vertices which support no propagators, while for $P \neq \emptyset$ we have \[F(P) = \big(V(P) \setminus V(P^c)\big) \cup F(\emptyset).\] \note{updated (1) to match new definition of $F(P)$; }
\item If $Q\subseteq P$ then $F(Q) \subseteq F(P)$.
\item $F(P)$ is a flat of $M(W)$, thus justifying the name ``propagator flat''.
\end{enumerate}
\end{lem} \sanote{I'd like to attach this to the "generic facts about matroids" in lemma 2.2?}
\begin{proof}
The proof of (1) and (2) are routine applications of the definition and are omitted.

@@

Let $v \in C$. Clearly $\Prop(v) \subseteq \Prop(C)$. Since $F(P)  = V(P^c)^c$, we have $F(\Prop(v)) \subseteq F(\Prop(C))$. Since $v \in F(\Prop(v))$ by the definition of propagator flat, we have $v \in F(\Prop(C))$ as required, and $C \subseteq F(\Prop(C))$.

Now suppose there exists some $w \in  F(\Prop(C)) \setminus C$. Let $B$ be an independent subset of $C$ of maximal rank; we first show that $B \cup \{w\}$ is a dependent set in $M(W)$. By Lemma \ref{lem facts about WLD matroids}, \bas |\Prop(C)| = \rk(C) = |B| = \rk(B) .\eas Since $B \subseteq C$ implies that  $\Prop(B) \subseteq \Prop(C)$, this implies that $\Prop(B) = \Prop(C)$. Furthermore, since $w \in F(Prop(C)$, $\Prop(B\cup w) \subseteq \Prop(B)$. By Lemma \ref{lem facts about WLD matroids} $\rk (B \cup w) \leq \min\{|B\cup w|, |\Prop(B\cup w)|\}$, and therefore $\rk (B \cup w) < |B\cup w|$. That is, $B\cup w$ is a circuit in $F(\Prop(C))$. Therefore, $B \cup w \subset C$, leading to a contradiction.

@@

For any $e \not\in F$, we certainly have $\rk((F\setminus C)\cup\{e\}) = \rk(F\setminus C) +1$ since $F$ is a flat. Now let $e \in C$, and suppose that $\rk((F\setminus C)\cup\{e\}) = \rk(F\setminus C)$. This implies that $(F\setminus C)\cup \{e\}$ is dependent, and hence contains a circuit, $S$. This circuit must contain at least one element but this contradicts the fact that $C$ was the union of all circuits in $F$.

@@

\begin{thm} \label{exactuniformthm}
Let $W':= (P, V(P))$ be a subdiagram of an admissible Wilson loop diagram $W= (\cP, [n])$. Then $W'$ is an exact subdiagram if and only if $M(W')$ is a uniform matroid of rank $|P|$.
\end{thm}

@@

    Taking all the possibilities for the left two cases of Figure~\ref{fig 3 cases}, the intervals $J^{(W)}_r(*)$ are still cyclic for all $r \neq p$, and we can assemble the intervals for the $J_p^{(W)}(*)$ as follows.
    \begin{itemize}
      \item 
    If $m=l$ then either $a=2$ (so $l+1=m+1$, $j=m+2$, and $j+1=m+3$) or $a=3$ (so $l+1=m+1$, $l+2=m+2$, $j=m+3$, and $j+1$ is not in the support of $p$), and in both cases
    \[
    J^{(W)}_p(m) = [m+4, m], \  J^{(W)}_p(m+1) = \{m+1\}, \  J^{(W)}_p(m+2) = \{m+2\}, \  J^{(W)}_p(m+3) = \{m+3\},
    \]
    which are nonempty and otherwise as required.
  \item
        If $m=i+1$ then checking each of the three different possibilities for $a$ we likewise get
    \[
    J^{(W)}_p(m) = [m+4, d-1], \  J^{(W)}_p(m+1) = [d, m+1], \  J^{(W)}_p(m+2) = \{m+2\}, \  J^{(W)}_p(m+3) = \{m+3\},
    \]
    which are nonempty and otherwise as required.
  \item If $m=i$ then $a=1$ or $a=2$, in the former case $l=m+2$, $j=m+3$ and $j+1$ is not in the support of $p$ so
    \[
    J^{(W)}_p(m) = \{m+4\}, \  J^{(W)}_p(m+1) = [m+5, d-1], \  J^{(W)}_p(m+2) = [d, m+2], \  J^{(W)}_p(m+3) = \{m+3\},
    \]
    while in the latter $l=m+2$, $l+1=m+3$, and $j$ and $j+1$ are not in the support of $p$ so
    \[
    J^{(W)}_p(m) = [m+4, j+1], \  J^{(W)}_p(m+1) = [j+2, d-1], \  J^{(W)}_p(m+2) = [d, m+2], \  J^{(W)}_p(m+3) = \{m+3\},
    \]
    which are again as required.
    \end{itemize}
